---
title: "Data Manipulation"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Load in the FAS litters Data

```{r}
litters_df = read_csv("./data/Fas_litters.csv")
litters_df = janitor :: clean_names(litters_df)
```

##  `select`

Choose some colums and not others by listing the name of those columns.
```{r}
select(litters_df, group, gd0_weight)
```

Select from one column to another by using ":".

```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```

Select to lose perticular column "litter_number" by using "-" .
```{r}
select(litters_df, -litter_number)
```

Renaming columns by using "select", result only shown the columns that changed. 

```{r}
select(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

Renaming columns by using "rename", renamed the columns but output the whole table.
```{r}
rename(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

select helpers

To select particular column: "starts_with" to select any columns that name "gd"
```{r}
select(litters_df, starts_with("gd"))
```

select "litter_number" and keep "everything"
```{r}
select(litters_df, litter_number, everything())
```

relocate column
```{r}
relocate (litters_df, litter_number)
```








