---
title: "Data Manipulation"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Load in the FAS litters Data

```{r}
litters_df = read_csv("./data/Fas_litters.csv")
litters_df = janitor :: clean_names(litters_df)
```

##  `select`

Choose some colums and not others by listing the name of those columns.
```{r}
select(litters_df, group, gd0_weight)
```

Select from one column to another by using ":".

```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```

Select to lose perticular column "litter_number" by using "-" .
```{r}
select(litters_df, -litter_number)
```

Renaming columns by using "select", result only shown the columns that changed. 

```{r}
select(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

Renaming columns by using "rename", renamed the columns but output the whole table.
```{r}
rename(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

select helpers

To select particular column: "starts_with" to select any columns that name "gd"
```{r}
select(litters_df, starts_with("gd"))
```

select "litter_number" and keep "everything"
```{r}
select(litters_df, litter_number, everything())
```

relocate column
```{r}
relocate (litters_df, litter_number)
```

## `filter`

filter gd0_weight < 22
```{r}
filter(litters_df, gd0_weight < 22)
```

filter gd0_weight >= 22
```{r}
filter(litters_df, gd0_weight >= 22)
```

"==" test for equality, asking if gd_of_birth = 20? 
```{r}
filter(litters_df, gd_of_birth == 20)
```

put"!" fliter anything is not = 20. Apply for more complicated situation. 
```{r}
filter(litters_df, !(gd_of_birth == 20))
```
OR , in simple case: 
```{r}
filter(litters_df, gd_of_birth != 20)
```
shown the same thing



```{r}
filter(litters_df, gd0_weight >=22, gd_of_birth == 20)
```


```{r}
filter(litters_df, group == "Mod8")
```

filter row that named "Con7" and "Mod8" 
```{r}
filter(litters_df, group %in% c("Con7", "Mod8"))
```

## `mutate`

this collection
1. create a new variable "wt_gain"
2. modify the existing variable "group" 


```{r}
litters_df <- litters_df %>%
  mutate(
    gd18_weight = as.numeric(gd18_weight),
    gd0_weight = as.numeric(gd0_weight),
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)
  )
```

## `arrange`

```{r}
arrange(litters_df, pups_born_alive)
```



















